% figures/case_study.tex
% Case study diagram - top to bottom forest
% Usage: \input{figures/case_study.tex}

\resizebox{\columnwidth}{!}{%
  \begin{forest}
    forked edges,
    for tree={
        grow'=south,           % top to bottom
        draw,
        rounded corners,
        font=\small,
        inner sep=3pt,
        minimum height=1.5em,
        node options={align=center},
        anchor=north,
        parent anchor=south,
        child anchor=north,
        l sep=1.2em,
        s sep=0.8em,
        edge={-latex},
      },
    % === PRIMITIVES ===
    % Rounded rectangle (data color - light blue): fill=clightblue
    % Rounded rectangle (infra color - light crest): fill=clightcrest
    % Rounded rectangle (readiness color - light purple): fill=clightpurple
    % Standard grey rectangle: fill=cslate1, sharp corners
    %
    % Example tree structure:
    % \begin{noindent}
    [{Rwanda's 60,000 CHW's lack reliable LLM based clinical decision support at national scale \citep{rutunda2026llmfrontline}}, fill=cslate1, sharp corners, text width=12em
      [{Working model can't be deployed?}, fill=clightcrest, text width=8em,
        [{CHWs lack devices or network?}, fill=clightcrest, text width=6em
          [{No}, fill=cslate1, sharp corners, text width=2em]
        ]
        [{Too expensive to deploy?}, fill=clightcrest, text width=4em
          [{\textbf{Yes}}, fill=cslate1, sharp corners, text width=2em
            [{Per-token API price too high?}, fill=clightcrest, text width=8em
              [{No}, fill=cslate1, sharp corners, text width=2em]
            ]
            [{Token price per query too high?}, fill=clightcrest, text width=8em
              [{\textbf{Yes}}, fill=cslate1, sharp corners, text width=2em
                [{Higher output token counts?}, fill=clightblue, text width=6em
                  [{\textbf{Yes}. Subword tokenization of Kinyarwanda produces more tokens \citep{ahia2023do}}, fill=cslate1, sharp corners, text width=14em
                    [{\textbf{Binding constaint: Data}}, fill=clightblue, sharp corners, text width=10em]
                  ]
                ]
                [{No language support?}, fill=clightblue, text width=8em]
              ]
            ]
          ]
        ]
        [{Cannot coordinate deployments}, fill=clightcrest, text width=6em
          [{No}, fill=cslate1, sharp corners, text width=2em]
        ]
      ]
      [{No models that perform well on clinical trials?}, fill=clightblue, text width=8em,
          [{No}, fill=cslate1, sharp corners, text width=2em]
      ]
    ]
    % \end{noindent}
    %  [{\textbf{Binding constaint: Data}}, fill=clightblue, sharp corners, text width=10em]
  \end{forest}%
}


% [{Higher output token counts on Kinyarwanda vs. English, driving up costs}, fill=clightblue, text width=12em]